---
title: "R Notebook"
output: html_notebook
---

Libraries needed

```{r}
library(dplyr)
library(GGally)
library(corrplot)
library(caret)
library(psych)
library(rpart)
library(randomForest)
library(nnet)
library(e1071)
library(naivebayes)

```

Loading the data


```{r}
filename <- '/Users/swapnilvermani/Documents/diabetic_data.csv'
raw_data <- read.table(filename, sep = ",", header = T, na.strings = "?")
head(raw_data)
summary(raw_data)
```

For getting the readmission prediction we dont need many columns hence we should be selecting only relevant values to make it simpler.Also we need to remove those columns which mainly has NA values.

```{r}
data <- select(raw_data,  -encounter_id, -patient_nbr, -weight,-(25:41),-(43:47))

head(data)

```
Basic summary of selected data and correlation plot of numeric data

```{r}
summary(data)
#ggpairs(data ,cardinality_threshold=NULL)
numeric_data<-select_if(data,is.numeric)
c <- cor(numeric_data, use= "pairwise.complete.obs")
corrplot(c)
```

```{r}
data$race[is.na(data$race)] <- "Other"

```

```{r}
plot(data$age, main = "age distribution") # age: mode 70-80yrs normal distribution, right skewed
plot(data$gender, main = "gender distribution") # gender: female 53% male 47%
plot(data$A1Cresult, main = "A1C") # A1Cresult: 84% no A1c results, 8% >8
plot(data$readmitted, main = "readmissions") # readmission: >50% no readmission
plot(data$admission_source, main = "admission source") # emergency 60%
```


Changing Categorical data

According to the glosssary provided and some extra help from wikipedia (https://en.wikipedia.org/wiki/List_of_ICD-9_codes) we can maniupulate the categorical variables in ordwer to have actual results

```{r}
#################################################################

# Changing categorical vriables

data2 <- data

data2$diag_1 <- as.numeric(levels(data2$diag_1)[data2$diag_1])
data2$diag_2 <- as.numeric(levels(data2$diag_2)[data2$diag_2])
data2$diag_3 <- as.numeric(levels(data2$diag_3)[data2$diag_3])

# diagnosis1
data2$diagnosis_group <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_group[data2$diag_1>=390 & data2$diag_1 <= 459 | data2$diag_1==785] <- "circulatory"
data2$diagnosis_group[data2$diag_1>=460 & data2$diag_1 <= 519 | data2$diag_1==786] <- "respiratory"
data2$diagnosis_group[data2$diag_1>=520 & data2$diag_1 <= 579 | data2$diag_1==787] <- "Digestive"
data2$diagnosis_group[data2$diag_1>=250 & data2$diag_1 < 251] <- "Diabetes"
data2$diagnosis_group[data2$diag_1>800 & data2$diag_1 <= 999] <- "Injury"
data2$diagnosis_group[data2$diag_1>=710 & data2$diag_1 <= 739] <- "Musculoskeletal"
data2$diagnosis_group[data2$diag_1>=580 & data2$diag_1 <= 629 | data2$diag_1==788] <- "Genitourinary"
data2$diagnosis_group[data2$diag_1>=140 & data2$diag_1 <= 239 | data2$diag_1>=790 & 
                        data2$diag_1 <= 799 | data2$diag_1==780 | data2$diag_1>=240 & data2$diag_1 < 250 |
                        data2$diag_1>=251 & data2$diag_1 <= 279 | data2$diag_1>=680 & data2$diag_1 <= 709 |
                        data2$diag_1>=001 & data2$diag_1 <= 139 | data2$diag_1==781 |
                      data2$diag_1==782 | data2$diag_1==784] <- "Neoplasms"

# diagnosis_2
data2$diagnosis_2 <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_2[data2$diag_2>=390 & data2$diag_2 <= 459 | data2$diag_2==785] <- "circulatory"
data2$diagnosis_2[data2$diag_2>=460 & data2$diag_2 <= 519 | data2$diag_2==786] <- "respiratory"
data2$diagnosis_2[data2$diag_2>=520 & data2$diag_2 <= 579 | data2$diag_2==787] <- "Digestive"
data2$diagnosis_2[data2$diag_2>=250 & data2$diag_2 < 251] <- "Diabetes"
data2$diagnosis_2[data2$diag_2>800 & data2$diag_2 <= 999] <- "Injury"
data2$diagnosis_2[data2$diag_2>=710 & data2$diag_2 <= 739] <- "Musculoskeletal"
data2$diagnosis_2[data2$diag_2>=580 & data2$diag_2 <= 629 | data2$diag_2==788] <- "Genitourinary"
data2$diagnosis_2[data2$diag_2>=140 & data2$diag_2 <= 239 | data2$diag_2>=790 & 
                        data2$diag_2 <= 799 | data2$diag_2==780 | data2$diag_2>=240 & data2$diag_2 < 250 |
                        data2$diag_2>=251 & data2$diag_2 <= 279 | data2$diag_2>=680 & data2$diag_2 <= 709 |
                        data2$diag_2>=001 & data2$diag_2 <= 139 | data2$diag_2==781 |
                        data2$diag_2==782 | data2$diag_2==784] <- "Neoplasms"

# diagnosis_3
data2$diagnosis_3 <- factor( rep("other",nrow(data2)),ordered = F, 
                                 levels = c("circulatory","respiratory","Digestive","Diabetes","Injury",
                                            "Musculoskeletal","Genitourinary","Neoplasms","other"))
data2$diagnosis_3[data2$diag_3>=390 & data2$diag_3 <= 459 | data2$diag_3==785] <- "circulatory"
data2$diagnosis_3[data2$diag_3>=460 & data2$diag_3 <= 519 | data2$diag_3==786] <- "respiratory"
data2$diagnosis_3[data2$diag_3>=520 & data2$diag_3 <= 579 | data2$diag_3==787] <- "Digestive"
data2$diagnosis_3[data2$diag_3>=250 & data2$diag_3 < 251] <- "Diabetes"
data2$diagnosis_3[data2$diag_3>800 & data2$diag_3 <= 999] <- "Injury"
data2$diagnosis_3[data2$diag_3>=710 & data2$diag_3 <= 739] <- "Musculoskeletal"
data2$diagnosis_3[data2$diag_3>=580 & data2$diag_3 <= 629 | data2$diag_3==788] <- "Genitourinary"
data2$diagnosis_3[data2$diag_3>=140 & data2$diag_3 <= 239 | data2$diag_3>=790 & 
                        data2$diag_3 <= 799 | data2$diag_3==780 | data2$diag_3>=240 & data2$diag_3 < 250 |
                        data2$diag_3>=251 & data2$diag_3 <= 279 | data2$diag_3>=680 & data2$diag_3 <= 709 |
                        data2$diag_3>=001 & data2$diag_3 <= 139 | data2$diag_3==781 |
                        data2$diag_3==782 | data2$diag_3==784] <- "Neoplasms"

# admission_source
data2$admission_source <- factor( rep("other",nrow(data2)),ordered = F, 
                             levels = c("clinic_referral", "emergency","other"))
data2$admission_source[data2$admission_source_id==c(1,2,3)]<- "clinic_referral"
data2$admission_source[data2$admission_source_id==7]<- "emergency"

# discharged_to
data2$discharged_to <- factor( rep("transferred",nrow(data2)),ordered = F, 
                                  levels = c("home", "transferred","left_AMA"))
data2$discharged_to[data2$discharge_disposition_id==c(1,6,8)]<- "home"
data2$discharged_to[data2$discharge_disposition_id==7]<- "left_AMA"

data2 <- select(data2, -diag_1, -diag_2, -diag_3, -admission_type_id, -discharge_disposition_id)
data2 <- select(data2, -medical_specialty)
data2 <- rename(data2, diag1 = diagnosis_group, diag2=diagnosis_2, diag3 = diagnosis_3)

# payer_code
data2$payer_code2 <- factor( rep("other",nrow(data2)),ordered = F, 
                               levels = c("other", "self_pay"))
data2$payer_code2[data2$payer_code=="SP"]<- "self_pay"
data2 <- select(data2, -payer_code)
data2 <- select(data2, -admission_source_id)
data2 <- rename(data2, payer_code=payer_code2)

head(data2)
write.csv(data2,"processed_diabetes.csv")
#######################################################
```

PCA Principal Component Analysis
In order to reduce 23 dimensions or knowing the significant impact of our reduction we will do PCA.


```{r}

###PCA section###
numeric_data<-scale(numeric_data)
pcaObj <- princomp(numeric_data, cor = TRUE, scores = TRUE, covmat = NULL)
summary(pcaObj)
print(pcaObj)
names(pcaObj)
plot(pcaObj)
pcaObj$loadings
biplot(pcaObj,cex=0.7)
final_data <- as.data.frame(pcaObj$scores)


```

According to PCA,
The columns that are of utmost importance are as follows:

age
discharged_to
time_in_hospital
num_lab_procedures
num_procedures
num_medications
number_outpatient
number_emergency
number_inpatient
number_diagnoses
insulin
change
diabetesMed
diag1
diag2
diag3
A1Cresult



Converting the dataset into training data and validation data.


```{r}

set.seed(123)
inTrain <- createDataPartition(y = data2$readmitted, p = .67,list = FALSE)
train <- data2[ inTrain,]
test <- data2[-inTrain,]
nrow(train)
nrow(test) 
train
test
```
###Regression###

Since the output variable readmitted is a categorical variable we need to find logistic regression instead of linear Regression

First by taking all the dimensions, in order to get the significance.

```{r}
train_nonbinary<-train
test_nonbinary<-test
train$readmitted<-ifelse(train$readmitted ==train$readmitted[1],0,1)
test$readmitted<-ifelse(test$readmitted ==test$readmitted[1],0,1)
test
logit_model <- glm(readmitted ~., data=train, family=binomial(link = 'logit'))
summary(logit_model)
plot(logit_model)
pred_logit <- predict(logit_model,test, type = "response")
#pred_logit

pred_logit <- ifelse(pred_logit > 0.5, 1, 0)
#pred_logit
result<-as.data.frame(table(pred_logit,test$readmitted))
#result
CorrectlyPredicted <- result[1,3]+result[4,3]
accuracy <-CorrectlyPredicted/nrow(test)
accuracy

senstivity_result<-result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result<-result[1,3]/(result[3,3]+result[1,3])
specificity_result
```
Adding dimension which are most significant and then doing logistic regression

```{r}

normal_fit <- glm(readmitted ~ race+age+discharged_to+time_in_hospital+
             num_lab_procedures+num_procedures+num_medications+number_outpatient+
             number_emergency+number_inpatient+number_diagnoses+
             insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
           data=train, family=binomial(link = 'logit'))
summary(normal_fit)
pred_logit <- predict(normal_fit,test, type = "response")
pred_logit

pred_logit <- ifelse(pred_logit > 0.5, 1, 0)
pred_logit
result<-as.data.frame(table(pred_logit,test$readmitted))
result
CorrectlyPredicted <- result[1,3]+result[4,3]
accuracy <-CorrectlyPredicted/nrow(test)
accuracy

senstivity_result<-result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result<-result[1,3]/(result[3,3]+result[1,3])
specificity_result

```


###Rpart Decision Tree ####
R part decision Trees Prediction

Prediction using three level category

Training

```{r}
rpart_tree <- rpart(formula = readmitted ~ age+discharged_to+time_in_hospital+
                      num_lab_procedures+num_procedures+num_medications+number_outpatient+
                      number_emergency+number_inpatient+number_diagnoses+
                      insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
                    data=train_nonbinary, method = 'class')
summary(rpart_tree)


```

Prediction 

```{r}
pred_tree <- predict(rpart_tree, test_nonbinary, type="class")

```

Performance

```{r}
table(predict(rpart_tree, test_nonbinary, type="class"), test_nonbinary$readmitted)
result<-as.data.frame(table(predict(rpart_tree, test_nonbinary, type="class"), test_nonbinary$readmitted))
result
confusionMatrix(pred_tree, test_nonbinary$readmitted)

prop.table(table(test_nonbinary$readmitted, pred_tree),1)
```

Prediction in binary format

Training

```{r}
train$readmitted<-ifelse(train$readmitted ==train$readmitted[1],0,1)
test$readmitted<-ifelse(test$readmitted ==test$readmitted[1],0,1)

rpart_tree <- rpart(formula = readmitted ~ age+discharged_to+time_in_hospital+
                      num_lab_procedures+num_procedures+num_medications+number_outpatient+
                      number_emergency+number_inpatient+number_diagnoses+
                      insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
                    data=train, method = 'class')
summary(rpart_tree)
```

Prediction

```{r}
pred_tree <- predict(rpart_tree, test, type="class")

```

Performance

```{r}
table(predict(rpart_tree, test, type="class"), test$readmitted)
result<-as.data.frame(table(predict(rpart_tree, test, type="class"), test$readmitted))
result

CorrectlyPredicted <- result[1,3]+result[4,3]
accuracy <-CorrectlyPredicted/nrow(test)
accuracy

senstivity_result<-result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result<-result[1,3]/(result[3,3]+result[1,3])
specificity_result

prop.table(table(test$readmitted, pred_tree),1)
```

###Random Forest###

Prediction using random_forest

Prediction using three level category

Training
```{r}
forest_model<-randomForest(formula=readmitted ~ age+discharged_to+time_in_hospital+
                       num_lab_procedures+num_procedures+num_medications+number_outpatient+
                       number_emergency+number_inpatient+number_diagnoses+
                       insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult,
                     data=train_nonbinary)
print(forest_model)
```

Prediction

```{r}
pred_forest <- predict(forest_model, test, type = "response")
table(test_nonbinary$readmitted, pred_forest)
prop.table(table(test_nonbinary$readmitted, pred_forest),1)


```

Performance Metrics

```{r}
confusionMatrix(test_nonbinary$readmitted, pred_forest)
```

Prediction using binary level

Training

```{r}
forest_model<-randomForest(formula=readmitted ~ age+discharged_to+time_in_hospital+
                       num_lab_procedures+num_procedures+num_medications+number_outpatient+
                       number_emergency+number_inpatient+number_diagnoses+
                       insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult,
                     data=train)
print(forest_model)
```


```{r}
pred_forest <- predict(forest_model, test, type = "response")
table(test$readmitted, pred_forest)
prop.table(table(test$readmitted, pred_forest),1)
```




###Nueral Networks###
Prediction using Nueral Networks

Prediction using three level category

Training

```{r}
nnet_model <- nnet(formula = readmitted ~ race+age+discharged_to+time_in_hospital+
                     num_lab_procedures+num_procedures+num_medications+number_outpatient+
                     number_emergency+number_inpatient+number_diagnoses+
                     insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
                   data=train_nonbinary, size = 10, maxit = 100)

```

Prediction
```{r}
pred_nnet <- predict(nnet_model, test_nonbinary, type = "class")
table(test_nonbinary$readmitted,pred_nnet)
prop.table(table(test_nonbinary$readmitted, pred_nnet),1)
```

Performance
```{r}
pred_nnet<-factor(pred_nnet)
confusionMatrix(pred_nnet,test_nonbinary$readmitted)
```

Prediction using binary level

Training

```{r}
nnet_model <- nnet(formula = readmitted ~ race+age+discharged_to+time_in_hospital+
                     num_lab_procedures+num_procedures+num_medications+number_outpatient+
                     number_emergency+number_inpatient+number_diagnoses+
                     insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
                   data=train, size = 10, maxit = 100)
```

Prediction
```{r}
pred_nnet_binary <- predict(nnet_model, test ,type="raw")
pred_nnet_binary <- ifelse(pred_nnet_binary > 0.5, 1,0)

```

Performance
```{r}
result<-as.data.frame(table(pred_nnet_binary,test$readmitted))
result
CorrectlyPredicted <- result[1,3]+result[4,3]
accuracy <-CorrectlyPredicted/nrow(test)
accuracy

senstivity_result<-result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result<-result[1,3]/(result[3,3]+result[1,3])
specificity_result
```
###SVM###
Prediction using SVM

Prediction using three level category

Training

```{r}
SVMmodel <- svm(readmitted ~ age+discharged_to+time_in_hospital+
                  num_lab_procedures+num_procedures+num_medications+number_outpatient+
                  number_emergency+number_inpatient+number_diagnoses+
                  insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult,
                data=train_nonbinary, kernel = "linear")
                  
print(SVMmodel)
summary(SVMmodel)
```

Prediction
```{r}
x <- select(test, -readmitted)
y <- select(test, readmitted)
pred <- predict(SVMmodel, x)
pred_svm <- pred

```

Performance

```{r}
prop.table(table(test$readmitted, pred_svm),1)
confusionMatrix(test$pred_readmit, pred_svm)
```

Prediction using binary level

Training
```{r}
SVMmodel <- svm(readmitted ~ age+discharged_to+time_in_hospital+
                  num_lab_procedures+num_procedures+num_medications+number_outpatient+
                  number_emergency+number_inpatient+number_diagnoses+
                  insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult,
                data=train,kernel = "linear",
                 type = "C-classification", cross = 10, cost = 0.01, gamma = 1000)
                  
print(SVMmodel)
summary(SVMmodel)
```


Prediction
```{r}
pred_SVM_binary<-predict(SVMmodel,test,, decision.values = F)
pred_SVM_binary <- ifelse(pred_SVM_binary > 0.5, 1,0)
```

Performance
```{r}
result<-as.data.frame(table(pred_SVM_binary,test$readmitted))
result
CorrectlyPredicted <- result[1,3]+result[4,3]
accuracy <-CorrectlyPredicted/nrow(test)
accuracy

senstivity_result<-result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result<-result[1,3]/(result[3,3]+result[1,3])
specificity_result
```


###Naive Bayes###
Prediction using nbayes

Prediction using three level category

Training

```{r}
nbayesmodel <- naiveBayes(readmitted ~ age+discharged_to+time_in_hospital+
                            num_lab_procedures+num_procedures+num_medications+number_outpatient+
                            number_emergency+number_inpatient+number_diagnoses+
                            insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
                          data = train_nonbinary)
```

Prediction

```{r}
pred <- predict(nbayesmodel, test_nonbinary, type = "class")
pred_nbayes <- pred

```

Performance
```{r}
prop.table(table(test_nonbinary$readmitted, pred_nbayes),1)
pred_nbayes<-factor(pred_nbayes)
confusionMatrix(pred_nbayes, test_nonbinary$readmitted)
```
Prediction using binary level

Training

```{r}
train$readmitted <- as.factor(train$readmitted)
test$readmitted <- as.factor(test$readmitted)
length(test$readmitted)
nbayesmodel_binary <- naive_bayes(readmitted ~ age+discharged_to+time_in_hospital+
                            num_lab_procedures+num_procedures+num_medications+number_outpatient+
                            number_emergency+number_inpatient+number_diagnoses+
                            insulin+change+diabetesMed+diag1+diag2+diag3+A1Cresult, 
                          data = train ,laplace = 1, usekernel = T, prior = NULL)
```

Prediction

```{r}
pred <- predict(nbayesmodel_binary, test, type = "prob")
pred_nbayes_binary <- pred

maxidx <- function(arr) {
        return(  which(arr == max(arr)) )
}
idx.nb <- apply(pred_nbayes_binary, c(1),maxidx)
```


Performance
```{r}
result<-as.data.frame(table(idx.nb-1,test$readmitted))
result
CorrectlyPredicted <- result[1,3]+result[4,3]
accuracy <-CorrectlyPredicted/nrow(test)
accuracy

senstivity_result<-result[4,3]/(result[2,3]+result[4,3])
senstivity_result

specificity_result<-result[1,3]/(result[3,3]+result[1,3])
specificity_result
```



